---
title: "p8105_hw5_jx2562"
author: "Jingyi"
date: "2023-11-13"
output: github_document
---


```{r, message=FALSE}
library(tidyverse)
library(purrr)
library(ggplot2)
```

## Problem 1

Load the homicides dataset. Categorized the raw data by cites and summarize the total number of homicides and the number of unsolved homicides

```{r}
homicide_df = read_csv("data/homicide-data.csv") |>
  mutate(city_state = paste(city, state, sep = ", ")) |>
  group_by(city_state) |>
  summarise(total_homicides = n(),
            unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest")))
```

Here we have a data frame which describe the total number of homicide in the cities and the number of unsolved homicides in the cities.

Now, we estimate the proportion of homicides that are unsolved in Baltimore, MD. 

```{r}
baltimore_data = homicide_df |>
  filter(city_state == "Baltimore, MD") 

baltimore_test = prop.test(pull(baltimore_data, unsolved_homicides), pull(baltimore_data, total_homicides)) |>
  broom::tidy()

print(paste("Estimated proportion for Baltimore: ", round(baltimore_test$estimate[1], 4)))
print(paste("Confidence interval for Baltimore: ", round(baltimore_test$conf.low, 4), "-", round(baltimore_test$conf.high, 4)))
```

Given the above pipeline, we can create a workflow to calculate the estimate proportions and CIs for each cities.

```{r}
run_prop_test <- function(city) {
  data <- homicide_df |>
    filter(city_state == city) 
  
  prop_test_result <- prop.test(pull(data, unsolved_homicides), pull(data, total_homicides))
  tidy_result <- broom::tidy(prop_test_result)
  
  tibble(
    estimated_proportion = tidy_result$estimate[1],
    conf_low = tidy_result$conf.low,
    conf_high = tidy_result$conf.high
  )
}
```

```{r, warning=FALSE}
unsolved_prop <- homicide_df |>
  mutate(prop_test_results = map(city_state, run_prop_test)) |>
  unnest()

head(unsolved_prop)
```

The `unsolved_prop` data frame present the estimated proportion and CIs for unsolved homicides in each city, along with the raw total count of homicides and unsolved homicides.

Now we draw a plot that shows the estimates and CIs for each city.
```{r}
unsolved_prop$city_state <- factor(unsolved_prop$city_state,  levels=unsolved_prop$city_state[order(unsolved_prop$estimated_proportion)])

ggplot(unsolved_prop, aes(x = city_state, y = estimated_proportion, ymin = conf_low, ymax = conf_high)) +
  geom_point(size = 2) +
  geom_errorbar(width = 0.2, position = position_dodge(0.2)) +
  labs(title = "Estimated Proportion of Unsolved Homicides with Confidence Intervals",
       x = "City",
       y = "Estimated Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

As the plot shows, **Tulsa, AL** shows the lowest estimated proportion of unsolved homicides case as well as the highest error bar. The small sample size may lead to the result. At the meantime, **Chicago, IL** shows the highest estimated proportion and a very narrow error bar, which may caused by a large sample size.


## Problem 2


